---
output:
  html_document: default
  pdf_document: default
---
#**Chapter 2: Simple Linear Regression**

##*Author: J Wooldridge*


###**Regression Practice**

```{r,echo=FALSE}
library(wooldridge)
library(dplyr)
library(ggplot2)
library(rsq) #For R sq
```

#**Exapmple 2.3 CEO Salary and Return on Equity**
```{r, echo=FALSE}
summary(ceosal1$salary)
summary(ceosal1$roe)

fit1=lm(salary~roe,ceosal1)
summary(fit1)

sal_roe_plot=ggplot(ceosal1,aes(x=roe,y=salary))+geom_smooth(method='lm',se=FALSE,col='dark green')+geom_point(col='green')+ggtitle("Scatterplot: CEO Salary vs ROE")+annotate("text", x=55, y=14500, label = "R^2 == 0.01", parse=T)+annotate("text", x=55, y=13000, label = "hat(alpha) == 963.2", parse=T)+annotate("text", x=55, y=11500, label = "hat(beta) == 18.5", parse=T)+theme_bw()
sal_roe_plot
```
  
**Interpretation: **  
If the ROE of the the firm increases by 1 percentage point than the CEO's salary increasese by **`r (round(fit1$coefficients[2],2))`** thousand dollars.
  
#**Exapmple 2.4 Wage and Education**
  
Consider the scatterplot of wage vs education from the wage dataset and the OLS model estimates
```{r,echo=FALSE}
#OLS
fit=lm(wage~educ,wage1)
summary(fit)
fit
#Scatterplot+Regression line
wage_edu_plot=ggplot(wage1,aes(x=educ,y=wage))+geom_point(col='cornflower blue')+ggtitle("Scatterplot: Wage vs Education")+geom_smooth(method='lm',se=FALSE,col='red')+annotate("text", x=2, y=24, label = "R^2 == 0.16", parse=T)+annotate("text", x=2, y=22, label = "hat(alpha) == -0.90", parse=T)+annotate("text", x=2, y=20, label = "hat(beta) == 0.54", parse=T)+theme_bw()
wage_edu_plot
```
  
**Interpretation: **
1 more year of education  increases wage by a half a dollar per hour. <span style="color: red;">NOTE</span>: The <span style="color: red;">intercept</span> value of <span style="color: red;">-.54</span> implies negative hourly wage for no education. This is absurd. A person who has no education will not be paying his employer for doing work.

#**Exapmple 2.5 Voting Outcomes and Campaign Expenditures**
  
Consider the scatterplot of wage vs education from the wage dataset and the OLS model estimates
```{r,echo=FALSE}
#OLS
head(vote1)
fit2=lm(voteA~shareA,vote1)
summary(fit2)
fit2
#Scatterplot+Regression line
vote_share_plot=ggplot(vote1,aes(x=shareA,y=voteA))+geom_point(col='salmon')+ggtitle("Scatterplot: Voting Outcomes vs Campaign Expenditure")+geom_smooth(method='lm',se=FALSE,col='red')+annotate("text", x=95, y=40, label = "R^2 == 0.85", parse=T)+annotate("text", x=95, y=35, label = "hat(alpha) == 26.81", parse=T)+annotate("text", x=95, y=30, label = "hat(beta) == 0.46", parse=T)+theme_bw()
vote_share_plot
```
    
**Interpretation: **
Vote share increases by almost half a percent if Cadidate A increases his campaign spending by 1 percent in the total campaign spending among both. 


#**Example 2.6 CEO Salary and ROE**
```{r,echo=FALSE}
#OLS

data=ceosal1

data2= data %>% select(roe,salary)%>%mutate(pred_salary=round(fitted.values(fit1),2),residual=round(fit1$residuals,2))



cov_r_roe=cov(data2$roe,data2$residual)
m_pred=mean(data2$pred_salary)
m_pred
m_sal=mean(data2$salary)
m_sal

summary(fit1)
cor_pred_sal_act_sal=round(cor(data2$pred_salary,data2$salary),2)
cor_pred_sal_act_sal^2

ggplot(data2,aes(residual))+geom_histogram(binwidth=250,fill='cornflower blue',col='black')+ggtitle("Histogram of residuals")+xlab('residuals')+theme_bw()+annotate("text",x=11000,y=46,label="Covariance(residual,roe)== 0",parse=T)+annotate("text",x=11000,y=41,label="Mean_Actaul_Salary==1281.12",parse=T)+annotate("text",x=11000,y=36,label="Mean_Fitted_Salary== 1281.12",parse=T)

```
  
#**Example 2.7 Wage and Education**   
No Analysis required
  
#**Example 2.8 CEO Salary and ROE**  
The correlation between predicted salary and observed salary is **`r cor_pred_sal_act_sal`**. This when squared yields **`r round(cor_pred_sal_act_sal^2,2)`**.
  
The results of regression of salary on roe are:  
```{r,echo=FALSE} 
summary(fit1)
```
  
This explains the plot
```{r,echo=FALSE} sal_roe_plot```
  
#**Example 2.9 Voting Outcomes and Campaign Expenditures**   
The regresssion outcome as indicated by the plot, cleary, seems to be a better fit for the model.
<span style='color:red;'> But, How much better?</span>
```{r,echo=FALSE}
vote_share_plot
```
  
We check the R-Square also called the, Coefficient of Determination. In this case the R square value is **`r  rsq(fit2)`**.
  
#**2-4 Units of Measurement and Functional Form (Page 36)**  
(1)Change in DEPENDENT VARIABLE UNIT  
Consider the case of CEO_Sal regressed on ROE. Here, CEO_Sal is interms of thousands of $ and ROE is measured as percentage. If CEO_Sal is instead changed to 100s, what would the impact on beta and alpha?

The new regression result is as follows:
```{r, echo=FALSE}
head(ceosal1)
data3= ceosal1 %>% select(salary,roe) %>% mutate(sal_100=salary*10) #Salary in 100 dollars
head(data3)

fit3=lm(sal_100~roe,data3)
fit3$coefficients
```
The old regression (CEO_sal on ROE) result was
```{r,echo=FALSE}
fit1$coefficients
```
Notice, both intercept and slope get multiplied by **10**.  
**In General, for every multiplication or division by any constant 'c' to the DEPENDENT variable must be repeated across all intercept and independent variable.**

(2)Change in INDEPENDENT VARIABLE UNIT  
Consider the case of CEO_Sal regressed on ROE. Here, CEO_Sal is interms of thousands of $ and ROE is measured as percentage. If ROE is instead changed to its decimal form, what would the impact on beta and alpha?

The new regression result is as follows:
```{r, echo=FALSE}
head(ceosal1)
data4= ceosal1 %>% select(salary,roe) %>% mutate(roe_dec=roe/100) #roe in decimal
head(data4)

fit4=lm(salary~roe_dec,data4)
fit4$coefficients
```
The old regression (CEO_sal on ROE) result was
```{r,echo=FALSE}
fit1$coefficients
```
Notice, only beta changes and alpha remains unchanged. ROE is multiplied by 100 for a division of 100 of the independent's units. 

**In General, every multiplication or division by any constant 'c' to the INDEPENDENT variable must be repeated with division or multiplication resp of across DEPENDENT ONLY.nexr**

