---
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Introductory Econometrics: Chapter 1
## Author: J M Woolridge
## R Code Compilation by RJ Neel


# End of Chapter 1 exercises (Page 39)

### Computer Exercises 

**C1: Use the data in WAGE1 for this exercise** 

```{r, results="hide"}
## Computer Exercise C1
library(wooldridge) #load the Woolridge Package
wage1
?wage1 #Description of the dataset
head(wage1) #First 6 rows. Easy to view
ncol(wage1) # No of rows
nrow(wage1) #No. of columns

library(dplyr)
glimpse(wage1)
```


**(i) Find the average education level in the sample. What are the lowest and highest years of education?** 

*Solution*
```{r}
summary(wage1$educ)
#Alternatively
mean(wage1$educ) #avg education level
min(wage1$educ) #min education leve
max(wage1$educ) #max
```

**(ii) Find the average hourly wage in the sample. Does it seem high or low? **

*Solution*
```{r}
mean(wage1$wage) #Gives you the average hourly wage 
summary(wage1$wage) #Wage appears to be low
hist(wage1$wage) #Clearly skewed towards right

library(ggplot2)
ggplot(wage1,aes(wage))+geom_histogram(fill='cornflowerblue',col='black')
```

**(iii) The wage data are reported in 1976 dollars. Using the Internet or a printed source, find the Consumer Price Index (CPI) for the years 1976 and 2013.**

*Solution*
Using Table B-60 in the 2004 Economic Report of the President, the CPI was 56.9 in 1976 and 233 in 2013. 

**(iv) Use the CPI values from part (iii) to find the average hourly wage in 2013 dollars. Now does the average hourly wage seem reasonable? **

*Solution*
To convert 1976 dollars into 2013 dollars, we use the ratio of the CPIs, which is 233/56.9 ≈ 4.09. Therefore, the average hourly wage in 2013 dollars is roughly 4.09(\$\5.90) \≈\ \$\24.13, which is a reasonable figure.


**(v) How many women are in the sample? How many men?**

*Solution*
```{r}
head(wage1)
#Notice the female column is a binary variable implying 1 for female and 0 for male requring us to proceed with 'dplyr'
library(dplyr)
w=nrow(wage1 %>% group_by(female) %>% filter(female=='1'))
w
m=nrow(wage1)-w
m
```
The no of females in the sample are **```r w```** and males are **```r m```**
End of Computer Exercise 1

**C2: Use the data in BWGHT to answer this question** 
```{r}
head(bwght)
?bwght
```

(i) How many women are in the sample, and how many report smoking during pregnancy?

*Solution*
```{r}
#Note: This data set contains all women
nrow(bwght) #No of women smoking
nrow(bwght %>% filter(cigs>'0')) #Women that report smoking during pregnancy
```
The number of women in the sample are **```r nrow(bwght)```** and **```r nrow(bwght %>% filter(cigs>'0'))```** report smoking during pregnancy

(ii) What is the average number of cigarettes smoked per day? Is the average a good measure of the “typical” woman in this case? Explain. 

*Solution*
```{r}
mean(bwght$cigs)
summary(bwght$cigs)
hist(bwght$cigs)
#Based on the histogram and range it appears to be a good measure.
```
The average number of cigarettes smoked during pregnancy is **```r mean(bwght$cigs)```**. This is not a good measure as only a few women smoke. 

(iii) Among women who smoked during pregnancy, what is the average number of cigarettes smoked per day? How does this compare with your answer from part (ii), and why?
```{r}
ss=bwght %>% filter(cigs>'0')
head(ss)
m_ss=mean(ss$cigs)
#It markedly differs from the privious average by about 11 units.
summary(ss$cigs)
ggplot(ss,aes(x=cigs))+geom_histogram(fill='salmon',col='red',bin=30)+xlab('Cigarettes smoke per day')
```

**C3 The data in MEAP01 are for the state of Michigan in the year 2001. Use these data to answer the following questions.**
```{r}
head(meap01)
```

**(i) Find the largest and smallest values of math4. Does the range make sense? Explain. **

*Solution*
```{r}
head(meap01)
summary(meap01$math4) # It makes sense as percentage is between 0 and 100

```

**(ii) How many schools have a perfect pass rate on the math test? What percentage is this of the total sample? **

*Solution*
```{r}
library(dplyr)
passrate100=nrow(meap01 %>% group_by(math4) %>% filter(math4=='100'))
passrate100
samplesize=nrow(meap01)
samplesize

percent_passrate=round((passrate100/samplesize)*100,2)
percent_passrate
```


**(iii) How many schools have math pass rates of exactly 50%? **

*Solution*
```{r}
library(dplyr)

nrow(meap01 %>% group_by(math4) %>% filter(math4=='50'))

```


**(iv) Compare the average pass rates for the math and reading scores. Which test is harder to pass? **

*Solution*
```{r}
pass_m=mean(meap01$math4)
pass_m

pass_r=mean(meap01$read4)
pass_r

#Cleary Reading is much more difficult to pass

```


**(v) Find the correlation between math4 and read4. What do you conclude? **

*Solution*
```{r}
cor(meap01$math4,meap01$read4)
library(ggplot2)
ggplot(data=meap01,aes(x=meap01$math4,y=meap01$read4))+geom_point(col='salmon')+ggtitle("Scatterplot: Maths vs Reading")+xlab("Math Score")+ylab("Reading Score")
#It is strongly positive
```


**(vi) The variable exppp is expenditure per pupil. Find the average of exppp along with its standard deviation. Would you say there is wide variation in per pupil spending? **
```{r}
mean(meap01$exppp)
sd(meap01$exppp)
summary(meap01$exppp)
library(ggplot2)
ggplot(data=meap01,aes(x=meap01$exppp))+geom_histogram(col='dark grey',fill='white')+ggtitle("Histogram: Expenditure per Pupil")+xlab("Expenditure")+ylab("Frequency")
#Considering Min=1207 and Max=11958, It is significantly wide
```


**(vii) Suppose School A spends $6,000 per student and School B spends $5,500 per student. By what percentage does School A’s spending exceed School B’s? Compare this to 100 · [log(6,000) – log(5,500)], which is the approximation percentage difference based on the difference in the natural logs. (See Section A.4 in Appendix A.)**

*Solution*
```{r}
round((log(6000)-log(5500))*100,2) # Gives the Percentage
round(((6000-5500)/ 5500)*100,2)

```

**C4: The data in JTRAIN2 come from a job training experiment conducted for low-income men during 1976–1977; see Lalonde (1986). **
```{r}
head(jtrain2)
?jtrain2
```

**(i) Use the indicator variable train to determine the fraction of men receiving job training.**

*Solution*
```{r}
tr=nrow(jtrain2 %>% group_by(train) %>% filter(train=='1'))
tr
total=nrow(jtrain2)
total
percent_trained=round((tr/total)*100,2)
percent_trained
```


**(ii) The variable re78 is earnings from 1978, measured in thousands of 1982 dollars. Find the averages of re78 for the sample of men receiving job training and the sample not receiving job training. Is the difference economically large? **

*Solution*
```{r}
train78=jtrain2 %>% select(re78,train) %>% group_by(re78) %>% filter(train=='1')
train78
mean(train78$re78)
untrain78=jtrain2 %>% select(re78,train) %>% group_by(re78) %>% filter(train=='0')
untrain78
mean(untrain78$re78)

library(sm)
ab=c(train78$re78,untrain78$re78)
hist(train78$re78)
hist(untrain78$re78)
g=rep(1:2, rep(100,2))
sm.density.compare(ab,g, xlab="Earnings, 1978",model="equal") # plot densities

#Mean for Trained worker is 6.35 and Untrained is 4.55. Significant Difference

```

**(iii) The variable unem78 is an indicator of whether a man is unemployed or not in 1978. What fraction of the men who received job training are unemployed? What about for men who did not receive job training? Comment on the difference. **

*Solution*
```{r}
train_uemp=jtrain2 %>% select(unem78,train) %>% group_by(train) %>% filter(train=='1' & unem78=='1')
tu=nrow(train_uemp)
tu
tot_train=nrow(jtrain2 %>% select(unem78,train) %>% group_by(train) %>% filter(train=='1'))
tot_train

round((tu/tot_train)*100,2) #Percentage Unemployed with Training

untrain_unemp=jtrain2 %>% select(unem78,train) %>% group_by(train) %>% filter(train=='0' & unem78=='1')
utue=nrow(untrain_unemp)
utue
tot_untrain=nrow(jtrain2 %>% select(unem78,train) %>% group_by(train) %>% filter(train=='0'))

round((utue/tot_untrain)*100,2) #Percentage Unemployed without Training
#Unemployment figure among untrained individuals is much larger than trained
```


**(iv) From parts (ii) and (iii), does it appear that the job training program was effective? What would make our conclusions more convincing?**
 
*Solution*
The differences in earnings and unemployment rates suggest the training program had
strong, positive effects. Our conclusions about economic significance would be stronger if we
could also establish statistical significance

*End of C4*

**C5 The data in FERTIL2 were collected on women living in the Republic of Botswana in 1988. The variable children refers to the number of living children. The variable electric is a binary indicator equal to one if the woman’s home has electricity, and zero if not. **

```{r}
head(fertil2)
?fertil2
```


**(i) Find the smallest and largest values of children in the sample. What is the average of children? **

*Solution*
```{r}
summary(fertil2$children)
```

**(ii) What percentage of women have electricity in the home? **

*Solution*
```{r}
library(dplyr)
tot_women=nrow(fertil2)
tot_women
elec_women=nrow(fertil2 %>% filter(electric=="1"))
elec_women
perc_with_elect=round((elec_women/tot_women)*100,2)
perc_with_elect
```


**(iii) Compute the average of children for those without electricity and do the same for those with electricity. Comment on what you find. **

*Solution*
```{r}
library(dplyr)
elect_child=fertil2 %>% select(children,electric) %>% filter(electric==1)

avg_chil_with_elec=mean(elect_child$children) 

avg_chil_with_elec #Average Children in Electricy households


non_elect_child=fertil2 %>% select(children,electric) %>% filter(electric==0)

avg_chil_with_no_elec=mean(non_elect_child$children) 

avg_chil_with_no_elec #Average Children in Non Electricy households

#Clearly the electricity households have fewer children
```



(iv) From part (iii), can you infer that having electricity “causes” women to have fewer children? Explain.

*Solution*
Not directly byt may be influency of electricity devices like TV and radio are more likely to promote fewer children and more liberal thoughts.

**C6 Use the data in COUNTYMURDERS to answer this question. Use only the year 1996. The variable murders is the number of murders reported in the county. The variable execs is the number of executions that took place of people sentenced to death in the given county. Most states in the United States have the death penalty, but several do not. **
```{r}
head(countymurders)
?countymurders
#This is a Panel dataset
```


**(i) How many counties are there in the data set? Of these, how many have zero murders? What percentage of counties have zero executions? (Remember, use only the 1996 data.) **

*Solution*
```{r}
library(dplyr)
county96=countymurders %>% select(countyid,year,murdrate,execs) %>% filter(year=="1996")
nrow(county96)
county96_0murd=countymurders %>% select(countyid,year,murdrate,execs) %>% filter(year=="1996" & murdrate=="0")
nrow(county96_0murd) #Counties with Zero murders

county96_0exec=countymurders %>% select(countyid,year,murdrate,execs) %>% filter(year=="1996" & execs=="0")
nrow(county96_0exec) #Counties with Zero executions
perc_county_0exe=round(nrow(county96_0exec)/nrow(county96),2)*100
perc_county_0exe #Percentage of Counties with zero exexutions

```


**(ii) What is the largest number of murders? What is the largest number of executions? Why is the average number of executions so small? **

*Solution*
```{r}
library(dplyr)
county96 %>% filter(murdrate==max(county96$murdrate))
max(county96$murdrate) #Maximum Murder rate
max(county96$execs) #Maximum Executions
mean(county96$execs) #The mean of execution is so small because many counties do not practice death penalty and many have no executions

```

**(iii) Compute the correlation coefficient between murders and execs and describe what you find.**

*Solution*
```{r}
cor(county96$murdrate,county96$execs)
library(ggplot2)
ggplot(data=county96,aes(x=murdrate,y=execs))+geom_point(col='red')+ggtitle("Scatterplot: Murder rates vs Executions")+xlab("Murders")+ylab("Executions")
```


**(iv) You should have computed a positive correlation in part (iii). Do you think that more executions cause more murders to occur? What might explain the positive correlation?**

*Solution*
It is a very weak positive correlation. We would expect the two to have negative relationship. Here the data is influenced by the fact that many states have outlawed death penalty

*End of C6*

**C7 The data set in ALCOHOL contains information on a sample of men in the United States. Two key variables are self-reported employment status and alcohol abuse (along with many other variables). The variables employ and abuse are both binary, or indicator, variables: they take on only the values zero and one. **
```{r}
head(alcohol)
?alcohol
#Cross sectional data set
```




**(i) What is percentage of the men in the sample report abusing alcohol? What is the employment rate? **

**Solution**
```{r}
ab=nrow(alcohol %>% filter(abuse=="1"))
tot=nrow(alcohol)

p_ab=(round((ab/tot)*100,2))
p_ab #Percentage of men reporting alcohol abuse

employed=nrow(alcohol %>% filter(status=="3"))
employed
employment_rate=(round((employed/tot)*100,2))
employment_rate #Employment rate in the sample
```

**(ii) Consider the group of men who abuse alcohol. What is the employment rate? **

**Solution**
```{r}
ab_tot=nrow(alcohol %>% filter(abuse=="1"))
ab_tot
ab_employed=nrow(alcohol %>% filter(abuse=='1' & status=="3"))
ab_employed
ab_employment_rate=(round((ab_employed/ab_tot)*100,2))
ab_employment_rate #Employment rate among alcohol abuses in the sample
```
**(iii) What is the employment rate for the group of men who do not abuse alcohol? **
```{r}
no_ab_tot=nrow(alcohol %>% filter(abuse=="0"))
no_ab_tot
no_ab_employed=nrow(alcohol %>% filter(abuse=='0' & status=="3"))
no_ab_employed
no_ab_employment_rate=(round((no_ab_employed/no_ab_tot)*100,2))
no_ab_employment_rate #Employment rate among alcohol abuses in the sample
```

(iv) Discuss the difference in your answers to parts (ii) and (iii). Does this allow you to conclude that alcohol abuse causes unemployment?

*Solution*
There is hardly a difference of around 3%. Not much significant to make any conclusion with reasonabale certainty.

**THE END**