
## Introductory Econometrics: Chapter 1
## Author: J M Woolridge
## R Code Compilation by RJ Neel


# End of Chapter 1 exercises (Page 39)

### Computer Exercises 

**C1: Use the data in WAGE1 for this exercise** 

```{r, results="hide"}
## Computer Exercise C1
library(wooldridge) #load the Woolridge Package
wage1
?wage1 #Description of the dataset
head(wage1) #First 6 rows. Easy to view
ncol(wage1) # No of rows
nrow(wage1) #No. of columns
```


**(i) Find the average education level in the sample. What are the lowest and highest years of education?** 

*Solution*
```{r}
summary(wage1$educ)
#Alternatively
mean(wage1$educ) #avg education level
min(wage1$educ) #min education leve
max(wage1$educ) #max
```

**(ii) Find the average hourly wage in the sample. Does it seem high or low? **

*Solution*
```{r}
mean(wage1$wage) #Gives you the average hourly wage 
summary(wage1$wage) #Wage appears to be low
hist(wage1$wage) #Clearly skewed towards right
```

**(iii) The wage data are reported in 1976 dollars. Using the Internet or a printed source, find the Consumer Price Index (CPI) for the years 1976 and 2013.**

*Solution*
Using Table B-60 in the 2004 Economic Report of the President, the CPI was 56.9 in 1976 and 233 in 2013. 

**(iv) Use the CPI values from part (iii) to find the average hourly wage in 2013 dollars. Now does the average hourly wage seem reasonable? **

*Solution*
To convert 1976 dollars into 2013 dollars, we use the ratio of the CPIs, which is 233/56.9 ≈ 4.09. Therefore, the average hourly wage in 2013 dollars is roughly 4.09(\$\5.90) \≈\ \$\24.13, which is a reasonable figure.


**(v) How many women are in the sample? How many men?**

*Solution*
```{r}
head(wage1)
#Notice the female column is a binary variable implying 1 for female and 0 for male requring us to proceed with 'dplyr'
library(dplyr)
w=nrow(wage1 %>% group_by(female) %>% filter(female=='1'))
w
m=nrow(wage1)-w
m
```

End of Computer Exercise 1

**C2: Use the data in BWGHT to answer this question** 

(i) How many women are in the sample, and how many report smoking during pregnancy?

*Solution*
```{r}
#Note: This data set contains all women
nrow(bwght) #No of women smoking

```


(ii) What is the average number of cigarettes smoked per day? Is the average a good measure of the “typical” woman in this case? Explain. 

*Solution*
```{r}
mean(bwght$cigs)
summary(bwght$cigs)
hist(bwght$cigs)
#Based on the histogram and range it appears to be a good measure.
```
(iii) Among women who smoked during pregnancy, what is the average number of cigarettes smoked per day? How does this compare with your answer from part (ii), and why?
```{r}
avg_all=mean(bwght$cigs)
avg_all

library(dplyr)
nrow(bwght %>% group_by(cigs) %>% filter(cigs=='0'))
s=(bwght %>% group_by(cigs) %>% filter(cigs>'0'))
avg_s=mean(s$cigs)
avg_s
hist(s$cigs, xlab='Pregnant Women Smokers')
#It markedly differs from the privious average by about 11 units.
```


